---
title: "Developing Patterns (2) - Writing a geometry-based  pattern"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developing Patterns (2) - Writing a geometry-based  pattern}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.width = 8,
fig.height = 6
)

library(ggplot2)
library(ggpattern)
```

```{r setup}
library(ggplot2)
library(ggpattern)
```


Introduction
------------------------------------------------------------------------------

This vignette shows how to:

* write a `create_pattern()` function of the correct signature for a geometry-based pattern
* Instruct `{ggpattern}` on where to find this user defined pattern

### Table of Contents

* [Write a function with the correct arguments](#write)
* [Instruct ggpattern on how to use the pattern functon](#instruct)
* [Example](#example)
* [Next steps](#next-steps)



<a name="write"></a> Write the geometry-based pattern function
------------------------------------------------------------------------------


All geometry-based pattern creation functions must:

1. Have the exact function signature: `function(params, boundary_df, aspect_ratio, legend)`
    * `params` - parameters from the geom (the aesthetics) e.g `pattern_fill`
    * `boundary_df` - data.frame containing polygon information i.e. The `polygon_df` format.
    * `aspect_ratio` - the best guess that `ggpattern` is able to make as to the
      aspect ratio of the **viewport** in which this pattern is being rendered.
    * `legend` logical value to indicate whether or not this function is being called
      to render a key legend or the in-place geom fill.
2. Return a grid grob object. This can be any valid grob including a `grid::grobTree()`


Parameters for this pattern:

* `pattern_shape` to indicate the character to use for plotting
* `pattern_size` size of the plotted character
* `pattern_fill` colour of the character
* `pattern_alpha` character transparence


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create a single character pointsGrob at the cenroid of the geom area
#'
#' @param params params/coords for a single element. named list or single row data.frame
#' @param boundary_df mask for the pattern rendering
#' @param aspect_ratio a aspect ratio of the plotting area.
#' @param legend is the pattern being created in the legend? default FALSE.
#'  Use this flag if you want different pattern drawing behaviour for the legend.
#'
#' @return grid grob object
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
centroid_dot_pattern <- function(params, boundary_df, aspect_ratio, legend) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Convert the simple `boundary_df` polygon information into a 
  # simple features polygon object i.e. {sf}
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  boundary_sf <- ggpattern::convert_polygon_df_to_polygon_sf(boundary_df)
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Now that we have the boundary as an {sf} object, we can use a simple
  # features' function to find the centroid
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  centroid    <- sf::st_centroid(boundary_sf)
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a single character at the cenroid
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  grid::pointsGrob(
    x    = centroid[1],
    y    = centroid[2],
    pch  = params$pattern_shape,
    size = unit(params$pattern_size, 'char'),
    gp   = grid::gpar(
      col = ggplot2::alpha(params$pattern_fill, params$pattern_alpha)
    )
  )
}
```



<a name="instruct"></a> Let `{ggpattern}` know that there's an external pattern function it can use
------------------------------------------------------------------------------

A global option (`ggpattern_geometry_funcs`) is a named list which contains 
geometry-based pattern creating functions to use outside of `ggpattern`.

The **name** used in this list corresponds to the `pattern` name used with 
the geom - in this case we will be using `pattern = 'centroid'`.


```{r}
options(ggpattern_geometry_funcs = list(centroid = centroid_dot_pattern))
```


<a name="example"></a> Use this `centroid` pattern
------------------------------------------------------------------------------

There is an included `pattern = 'none'` in `ggpattern`.

Here, a verbose version of the 'none' pattern is created.  It still just returns
an empty grob, but as a side-effect prints out the parameters set for this object.

```{r}
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)


ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill          = trt,
      pattern_shape = trt,
      pattern_fill  = trt
    ),
    colour          = 'black',
    pattern         = 'centroid',
    pattern_size    = 3
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'centroid'"
  ) +
  scale_pattern_fill_viridis_d() + 
  theme(legend.key.size = unit(2, 'cm')) +
  coord_fixed(ratio = 1/2)
```





<a name="next-steps"></a>Next Steps
------------------------------------------------------------------------------

* Vignette [Developing Patterns (2) - Writing a geometry-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-2.html)
* Vignette [Developing Patterns (3) - Writing an array-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-3.html)
* `ggpattern` also includes some experimental patterns of varying completeness and 
   complexity:
    * [hexagons](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-hex-1.html) - geometry-based pattern
    * [ambient noise](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-ambient.html) - array-based pattern
    * [points](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-points.html) - geometry-based pattern
