---
title: "Developing Patterns (1) - Pattern Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developing Patterns (1) - Pattern Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
library(ggpattern)
```


Developing Patterns - overview
------------------------------------------------------------------------------

To develop a pattern for use with `ggpattern` you will need to:

1. Decide whether this is a geometry-based pattern or an array-based pattern.
2. Create a function with the required arguments and return value
3. Use `options()` to let `ggpattern` know that a certain pattern name should be
   delegated to your function.

### Table of Contents

* [The two pattern classes](#classes)
* [List of all pattern aesthetics](#aes)
* [Aesthetic use by pattern](#aes-by-pattern)
* [Geometry-based pattern functions](#geometry-based)
* [Array-based pattern functions](#array-based)
* [The `polygon_df` data.frame format](#polygon_df)
* [Specifying user pattern functions for ggpattern](#user-funcs)
* [Next steps](#next-steps)





<a name="classes"></a> Pattern Classification - geometry-based and array-based
------------------------------------------------------------------------------

There are only 2 high-level classes of pattern supported by `ggpattern` - geometry-based
and array-based.

**Geometry-based** patterns create a series of geometry objects and trim
them (using `sf:st_intersection()`, `polyclipGrob` etc) to be within the boundary
of the geom. For example, the 'stripes'
pattern in `ggpattern` is a series of equally spaced rectangular polygons.

**Array-based** patterns are RGBA image arrays.  Any supplied image will be 
processed by `ggpattern` to ensure it is masked to only apply to the area 
within the geom's boundary.



<a name="aes"></a> Aesthetics - Descriptions and Defaults
------------------------------------------------------------------------------


<details open = "true">
<summary> Aesthetic summary - click to open/close </summary>

| aesthetic                | description                                   | default    | possible values                    |
|--------------------------|-----------------------------------------------|------------|------------------------------------|
| `pattern`                  | Name of the pattern to draw                   | 'stripe'   | `gridpattern::names_pattern`       |
| `pattern_type`             | Generic control option                        | NA         | pattern-dependent                  |
| `pattern_subtype`          | Generic control option                        | NA         | pattern-dependent                  |
| `pattern_alpha`            | Alpha                                         | 1          | value in range [0, 1] or `NA`      |
| `pattern_linetype`         | Stroke linetype                               | 1          | linetype                           |
| `pattern_size`             | Stroke linewidth                              | 1          | linewidth                          |
| `pattern_shape`            | Plotting shape                                | 1          | shapes                             |
| `pattern_colour`           | Stroke colour                                 | 'grey20'   | colour                             |
| `pattern_fill`             | Fill colour                                   | 'grey80'   | colour                             |
| `pattern_fill2`            | Second fill colour                            | '#4169E1'  | colour                             |
| `pattern_angle`            | Rotation angle                                | 30         | angle in degrees                   |
| `pattern_density`          | Approx. fraction of area the pattern fills    | 0.2        | value in range [0, 1]              |
| `pattern_spacing`          | Spacing between repetitions of pattern        | 0.05       | value in range [0, 1] (snpc units) |
| `pattern_xoffset`          | Shift pattern along x axis                    | 0          | value in range [0, 1] (snpc units) |
| `pattern_yoffset`          | Shift pattern along y axis                    | 0          | value in range [0, 1] (snpc units) |
| `pattern_aspect_ratio`     | Aspect ratio adjustment                       | NA         | usual range [0.01, 10]             |
| `pattern_key_scale_factor` | Scale factor for pattern in legend            | 1          |                                    |
| `pattern_filename`         | Image filename/URL                            | ''         | Filename/URL                       |
| `pattern_gravity`          | Image placement                               | 'center'   | `magick::gravity_types`            |
| `pattern_filter`           | Image scaling filter                          | 'lanczos'  | `magick::filter_types`             |
| `pattern_scale`            | Scale                                         | 1          | Multiplier                         |
| `pattern_orientation`      | Orientation                                   | 'vertical' | 'vertical', 'horizontal', 'radial' |
| `pattern_phase`            | Phase                                         | 0          |                                    |
| `pattern_frequency`        | Frequency                                     | 0.1        |                                    |
| `pattern_option_1 - 5`     | Generic options for expansion                 | 0          |                                    |

</summary>
</details>

<a name="aes-by-pattern"></a> Aesthetics Usage by Pattern 
------------------------------------------------------------------------------

<details open = "true">
<summary> Aesthetic Usage-by-Pattern - click to open/close </summary>

| aesthetic                | circle | pch | polygon_tiling | regular_polygon |
|--------------------------|----------------|-------|-------------|-------------|
| pattern                  | Y              |  Y    |  Y          |  Y          |
| pattern_type             | Y              |  Y    |  Y          |  Y          |
| pattern_subtype          | Y              |  Y    |             |  Y          |
| pattern_alpha            | Y              |  Y    |             |  Y          |
| pattern_linetype         | Y              |  Y    |  Y          |  Y          |
| pattern_size             | Y              |  Y    |  Y          |  Y          |
| pattern_shape            |                |  Y    |             |  Y          |
| pattern_colour           | Y              |  Y    |  Y          |  Y          |
| pattern_fill             | Y              |  Y    |  Y          |  Y          |
| pattern_fill2            |                |       |             |             |
| pattern_angle            | Y              |  Y    |  Y          |  Y          |
| pattern_density          | Y              |  Y    |             |  Y          |
| pattern_spacing          | Y              |  Y    |  Y          |  Y          |
| pattern_xoffset          | Y              |  Y    |  Y          |  Y          |
| pattern_yoffset          | Y              |  Y    |  Y          |  Y          |
| pattern_aspect_ratio     | Y              |  Y    |  Y          |  Y          |
| pattern_key_scale_factor | Y              |  Y    |  Y          |  Y          |
| pattern_filename         |                |       |             |             |
| pattern_gravity          |                |       |             |             |
| pattern_filter           |                |       |             |             |
| pattern_scale            |                |       |             |             |
| pattern_orientation      |                |       |             |             |
| pattern_phase            |                |       |             |             |
| pattern_frequency        |                |       |             |             |
| pattern_option_1 - 5     |                |       |             |             |

| aesthetic                | stripe  | wave | crosshatch | weave |
|--------------------------|---------|------|------------|-------|
| pattern                  | Y       |  Y   |     Y      |   Y   |
| pattern_type             |         |  Y   |            |   Y   |
| pattern_subtype          |         |      |            |   Y   |
| pattern_alpha            | Y       |  Y   |     Y      |   Y   |
| pattern_linetype         | Y       |  Y   |     Y      |   Y   |
| pattern_size             | Y       |  Y   |     Y      |   Y   |
| pattern_shape            |         |      |            |       |
| pattern_colour           | Y       |  Y   |     Y      |   Y   |
| pattern_fill             | Y       |  Y   |     Y      |   Y   |
| pattern_fill2            |         |      |     Y      |   Y   |
| pattern_angle            | Y       |  Y   |     Y      |   Y   |
| pattern_density          | Y       |  Y   |     Y      |   Y   |
| pattern_spacing          | Y       |  Y   |     Y      |   Y   |
| pattern_xoffset          | Y       |  Y   |     Y      |   Y   |
| pattern_yoffset          | Y       |  Y   |     Y      |   Y   |
| pattern_aspect_ratio     | Y       |  Y   |     Y      |   Y   |
| pattern_key_scale_factor | Y       |  Y   |     Y      |   Y   |
| pattern_filename         |         |      |            |       |
| pattern_gravity          |         |      |            |       |
| pattern_filter           |         |      |            |       |
| pattern_scale            |         |      |            |       |
| pattern_orientation      |         |      |            |       |
| pattern_phase            |         |      |            |       |
| pattern_frequency        |         |  Y   |            |       |
| pattern_option_1 - 5     |         |      |            |       |

| aesthetic                | ambient | image | placeholder | magick | gradient | plasma |
|--------------------------|-------|-------|-------------|--------|----------|--------|
| pattern                  | Y     | Y     | Y           | Y      | Y        | Y      |
| pattern_type             | Y     | Y     | Y           |        |          | Y      |
| pattern_subtype          |       |       |             |        |          |        |
| pattern_alpha            | Y     | Y     | Y           | Y      | Y        | Y      |
| pattern_linetype         |       |       |             |        |          |        |
| pattern_size             |       |       |             |        |          |        |
| pattern_shape            |       |       |             |        |          |        |
| pattern_colour           |       |       |             |        |          |        |
| pattern_fill             | Y     |       |             | Y      | Y        | Y      |
| pattern_fill2            | Y     |       |             |        | Y        |        |
| pattern_angle            |       |       |             |        |          |        |
| pattern_density          |       |       |             |        |          |        |
| pattern_spacing          |       |       |             |        |          |        |
| pattern_xoffset          |       |       |             |        |          |        |
| pattern_yoffset          |       |       |             |        |          |        |
| pattern_aspect_ratio     | Y     | Y     | Y           | Y      |          | Y      |
| pattern_key_scale_factor | Y     | Y     | Y           | Y      | Y        | Y      |
| pattern_filename         |       | Y     |             |        |          |        |
| pattern_gravity          |       | Y     |             |        |          |        |
| pattern_filter           |       | Y     |             | Y      |          |        |
| pattern_scale            |       | Y     |             | Y      |          | Y      |
| pattern_orientation      |       |       |             |        | Y        |        |
| pattern_phase            |       |       |             |        |          |        |
| pattern_frequency        |       |       |             |        |          |        |
| pattern_option_1 - 5     |       |       |             |        |          |        |

There are a number of  aesthetics which are part of the package, but not currently used in any patterns: `pattern_subtype`, 
`pattern_phase`, `pattern_frequency` and `pattern_option_1` through to `pattern_option_5`.  These may be useful for
developing patterns outside of the package.

</summary>
</details>


<a name="geometry-based"></a> Geometry-based pattern functions - Formal Arguments and Return Values
------------------------------------------------------------------------------

All geometry-based pattern creation functions must:

1. Have the exact function signature: `function(params, boundary_df, aspect_ratio, legend)`
    * `params` - parameters from the geom (the aesthetics) e.g `pattern_fill`
    * `boundary_df` - data.frame containing polygon information i.e. The `polygon_df` format.
    * `aspect_ratio` - the best guess that `ggpattern` is able to make as to the
      aspect ratio of the **viewport** in which this pattern is being rendered.
    * `legend` logical value to indicate whether or not this function is being called
      to render a key legend or the in-place geom fill.
2. Return a grid grob object. This can be any valid grob including a `grid::grobTree()`


List of geometry-based patterns in `ggpattern`:

* circle
* crosshatch
* pch
* polygon_tiling
* regular_polygon
* stripe
* weave


<a name="array-based"></a> Array-based pattern functions - Formal Arguments and Return Values
------------------------------------------------------------------------------

All array-based pattern creation functions must:

1. Have the exact function signature: `function(width, height, params, legend)`
    * `width,height` - dimensions of the bounding box of the geom area
    * `params` - parameters from the geom (the aesthetics) e.g `pattern_fill`
    * `legend` logical value to indicate whether or not this function is being called
      to render a key legend or the in-place geom fill.
2. Return a 3D array of RGBA values (all values in the range [0, 1]).


List of array-based patterns in `ggpattern`:

* ambient
* gradient
* image
* magick
* placeholder
* plasma


<a name="polygon_df"></a> The `polygon_df` data.frame format
------------------------------------------------------------------------------

The `polygon_df` is a very simple data.frame format to contain polygon values.  This is
used to pass the coordinates of the geom boundary from the geom to the
pattern generating function.

It contains only 'x' and 'y' columns for the coordinates, and an 'id' column 
used to signify which polygon the coordinates belong to.

The following `polygon_df` data.frame contains 2 polygons:

```{r echo = FALSE}
x <- readr::read_csv("
x,  y,  id
0,  0,   1
1,  0,   1
1,  1,   1
0,  1,   1
0,  0,   2
2,  0,   2
2,  1,   2
0,  1,   2
")

knitr::kable(x, caption = "example data in 'polygon_df' format")
```


Limitations of the 'polygon_df' format:

* There is no support for polygons with holes.



<a name="user-funcs"></a> Associating a function with  `ggpattern` pattern name
------------------------------------------------------------------------------

There are two global `option()` values which can be set - one for geometry-based 
patterns, and the other for array-based patterns.

The global values should point to a named list, where the names are the pattern 
names you want to use within `ggpattern`, and the the named values are the
actual functions.


```{r eval = FALSE}
options(ggpattern_array_funcs    = list(your_pattern_name = your_pattern_function))
options(ggpattern_geometry_funcs = list(your_pattern_name = your_pattern_function))
```





<a name="next-steps"></a>Next Steps
------------------------------------------------------------------------------

* Vignette [Developing Patterns (2) - Writing a geometry-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-2.html)
* Vignette [Developing Patterns (3) - Writing an array-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-3.html)
* `ggpattern` also includes some experimental patterns of varying completeness and 
   complexity:
    * [hexagons](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-hex-1.html) - geometry-based pattern
    * [ambient noise](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-ambient.html) - array-based pattern
    * [points](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-points.html) - geometry-based pattern










