
---
title: "Developing Patterns (3) - Writing an array-based  pattern"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Developing Patterns (3) - Writing an array-based  pattern}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.width = 8,
fig.height = 6
)

library(ggplot2)
library(ggpattern)
```

```{r setup}
library(ggplot2)
library(ggpattern)
```


Introduction
------------------------------------------------------------------------------

This vignette shows how to:

* write a `create_pattern()` function of the correct signature for a array-based pattern
* Instruct `{ggpattern}` on where to find this user defined pattern

### Table of Contents

* [Write a function with the correct arguments](#write)
* [Instruct ggpattern on how to use the pattern functon](#instruct)
* [Example](#example)
* [Next steps](#next-steps)



<a name="write"></a> Write the array-based pattern function
------------------------------------------------------------------------------


All array-based pattern creation functions must:

1. Have the exact function signature: `function(width, height, params, legend)`
    * `width,height` - dimensions of the bounding box of the geom area
    * `params` - parameters from the geom (the aesthetics) e.g `pattern_fill`
    * `legend` logical value to indicate whether or not this function is being called
      to render a key legend or the in-place geom fill.
2. Return a 3D array of RGBA values (all values in the range [0, 1]).


Parameters for this pattern:

* `pattern_type` is used to distinguish between source data for the pattern 


**Note**: This pattern exploits vector recyling in the creation of the RGBA array, 
and as dimensions change the alignment of the R, G, B and A planes will not 
remain in a fixed relationship.  Thus if you change the shape of the rendered image,
you will change the nature of the pattern.

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create an array of texture
#'
#' @param width,height area dimensions
#' @param params aesthetic parameters passed from the geom e.g. 'pattern_fill', 
#'        'pattern_frequency' etc.
#' @param legend logical. If the request to create a pattern comes during 
#'        creation of the legend, then this is TRUE, otherwise FALSE
#'
#' @return an RGBA numeric array with dimensions [height, width, 4]
#'
#' @import ambient
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
create_pattern_simple <- function(width, height, params, legend) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Ensure the selected pattern is sane.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  choice <- params$pattern_type
  if (is.null(choice) || is.na(choice) || !is.character(choice)) {
    choice <- 'a'
  }
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Choose the values with which to fill the array
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  values <- switch(
    choice,
    a = rep(c(0, 1, 0, 1, 1, 0, 0, 1, 1, 1), each = 3),
    b = rep(c(1, 0, 0, 1, 0.5, 0.5, 1, 1, 0, 0, 0, 0, 0, 0.5), each = 7),
    c = rep(seq(0, 1, 0.05), each = 7),
    rep(c(0, 1, 0, 1, 1, 0, 0, 1, 1, 1), each = 3)
  )
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create an RGBA array of the requested dimensions
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  simple_array <- array(values, dim = c(height, width, 4))
  
  simple_array
}
```



<a name="instruct"></a> Let `{ggpattern}` know that there's an external pattern function it can use
------------------------------------------------------------------------------

A global option (`ggpattern_geometry_funcs`) is a named list which contains 
geometry-based pattern creating functions to use outside of `ggpattern`.

The **name** used in this list corresponds to the `pattern` name used with 
the geom - in this case we will be using `pattern = 'simple'`.


```{r}
options(ggpattern_array_funcs = list(simple = create_pattern_simple))
```


<a name="example"></a> Use this `simple` pattern
------------------------------------------------------------------------------


```{r}
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)


ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill         = trt,
      pattern_type = trt
    ),
    colour          = 'black',
    pattern         = 'simple',
    pattern_size    = 3
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'simple'"
  ) +
  scale_pattern_type_discrete(choices = c('a', 'b', 'c')) + 
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```


<a name="example"></a> Use this `simple` pattern
------------------------------------------------------------------------------

```{r}
ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x            = mpg, 
      pattern_type = as.factor(cyl)
    ), 
    fill    = 'white', 
    pattern = 'simple'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'simple'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm')) +
  scale_pattern_type_discrete(choices = c('a', 'b', 'c')) + 
  theme(legend.position = 'none') +
  coord_fixed(ratio = 100)
```


<a name="next-steps"></a>Next Steps
------------------------------------------------------------------------------

* Vignette [Developing Patterns (2) - Writing a geometry-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-2.html)
* Vignette [Developing Patterns (3) - Writing an array-based  pattern](https://coolbutuseless.github.io/package/ggpattern/articles/developing-patterns-3.html)
* `ggpattern` also includes some experimental patterns of varying completeness and 
   complexity:
    * [hexagons](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-hex-1.html) - geometry-based pattern
    * [ambient noise](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-ambient.html) - array-based pattern
    * [points](https://coolbutuseless.github.io/package/ggpattern/articles/experimental-pattern-points.html) - geometry-based pattern
