---
title: "Experimental Pattern - Hex"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Pattern - Hex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 6,
  fig.width = 8
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(sf)
  library(grid)
  library(ggpattern)
})
```


Introduction to the geometry-based `hex` pattern
------------------------------------------------------------------------------


The pattern is an attempt to create a structured pattern of hex elements using
only geometry elements.


Create the `hex` pattern function
------------------------------------------------------------------------------

All geometry-based pattern creation functions must:

1. Have the exact function signature: `function(params, boundary_df, aspect_ratio, legend)`
2. Return a grid grob object



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create an array of noise using the 'ambient' package
#'
#' @param params aesthetic parameters passed from the geom e.g. 'pattern_fill', 
#'        'pattern_frequency' etc.
#' @param boundary_df is a data.frame of (x, y) coordinates of the boundary of
#"        the geom to be filled.
#' @param aspect_ratio this is the best guess of the current aspect ratio of the
#'        viewport into which the geometry is being drawn
#' @param legend logical. If the request to create a pattern comes during 
#'        creation of the legend, then this is TRUE, otherwise FALSE
#'
#' @return a grid grob object containing the pattern
#'
#' @import ambient
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
create_pattern_hex <- function(params, boundary_df, aspect_ratio,
                               legend = FALSE) {

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Pre-scale + rotate the boundary boundary
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  boundary_df$y <- boundary_df$y /  aspect_ratio
  boundary_df   <- rotate_polygon_df(boundary_df, params$pattern_angle, aspect_ratio)


  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # 1. Convert the geometry boundary into an 'sf' object
  # 2. Make a hexagonal grid using `sf::st_make_grid()`
  # 3. Keep only the hexes within the boundary
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  boundary_sf        <- convert_polygon_df_to_polygon_sf(boundary_df)
  hex_sfc            <- st_make_grid(boundary_sf, cellsize = params$pattern_spacing, square = FALSE, flat_topped = FALSE)
  hex_within_area_sf <- st_intersection(hex_sfc, boundary_sf)

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Convert the multipolygon sf object containing hexes into a standard
  # polygon_df format.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  polys <- lapply(
    seq_along(hex_within_area_sf),
    function(idx) {
      mat <- as.matrix(hex_within_area_sf[[idx]])
      mat <- head(mat, -1)
      cbind(mat, idx)
    }
  )
  polys <- do.call(rbind, polys)
  polys <- as.data.frame(polys)
  polys <- setNames(polys, c('x', 'y', 'id'))

  hex_within_area_df   <- polys


  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Calculate the linewidth to simulate increases in density.
  # convert the 'npc' cellsize into native coordinates, and then set the
  # linewidth to be a fraction of that size
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  cell_size <- as.numeric(grid::convertWidth(unit(params$pattern_spacing, 'npc'), 'native'))
  lwd <- params$pattern_density * cell_size

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Rotate the pattern into its final position
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  hex_within_area_df <- rotate_polygon_df(hex_within_area_df, -params$pattern_angle, aspect_ratio)
  hex_within_area_df$y <- hex_within_area_df$y * aspect_ratio

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # return early if there are no hexes in the area
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (is.null(hex_within_area_df) || nrow(hex_within_area_df) == 0) {
    return(grid::nullGrob())
  }

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a polygonGrob all all the hexes and clipped hexes
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  hex_grob <- grid::polygonGrob(
    x = unit(hex_within_area_df$x, "npc"),
    y = unit(hex_within_area_df$y, "npc"),
    id = hex_within_area_df$id,
    gp = gpar(
      col     = scales::alpha(params$pattern_fill, params$pattern_alpha),
      fill    = NA,
      lwd     = lwd,
      lty     = params$pattern_linetype,
      lineend = 'square'
    )
  )

  hex_grob
}

```


Let `{ggpattern}` know that there's an external pattern function it can use
------------------------------------------------------------------------------

A global option (`ggpattern_geometry_funcs`) is a named list which contains 
grid creating functions to use outside of `ggpattern`.

The **name** used in this list corresponds to the `pattern` name used with 
the geom - in this case we will be using `pattern = 'hex'`.


```{r}
options(ggpattern_geometry_funcs = list(hex = create_pattern_hex))
```


Use this `hex` pattern
------------------------------------------------------------------------------


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create some data to plot
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a ggplot using this pattern
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
p <- ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_fill  = trt,
      pattern_angle = trt
    ),
    pattern         = 'hex', 
    pattern_spacing = 0.02, 
    pattern_density = 0.1,
    pattern_alpha   = 1, 
    fill            = 'white', 
    colour          = NA
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'hex'"
  ) +
  theme(legend.position = 'none') +
  scale_pattern_angle_discrete(range = c(0, 30)) + 
  coord_fixed(ratio = 1/2)

p
```


Use this `hex` pattern - thicker lines
------------------------------------------------------------------------------

Because of the way `sf::st_intersection()` works, the boundary of the geom gets
inextricably linked into the pattern.  

Unlike all other patterns, this means that as the pattern changes, the representation
of the *boundary* also changes.

In this example, the thicker lines for the internal hexes, also mean thicker lines
for the rectangular boundary.

I don't really like how this is different from all other patterns.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a ggplot using this pattern
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
p <- ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(pattern_fill = trt),
    pattern         = 'hex', 
    pattern_spacing = 0.02, 
    pattern_angle   = 0, 
    pattern_density = 0.5,
    pattern_alpha   = 1, 
    fill            = 'white', 
    colour          = NA
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'hex'"
  ) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)

p
```




Future
------------------------------------------------------------------------------

* intersection with border looks terrible at large stroke width.  If that problem
  could be worked out, then this would be a useful addition to the package.
* when rendered as small hexes, there isn't a significant advantage over 
 `pattern = 'magick'` with `pattern_type = 'hexagons'`

