
---
title: "Experimental Pattern - Noise with Ambient"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Pattern - Noise with Ambient}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.width = 9,
fig.height = 6
)
```

```{r setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(ggpattern)
  
  library(ambient)
})
```


Introduction to the array-based `ambient` pattern
------------------------------------------------------------------------------

The `ambient` package offers a way of creating multiple different types of 
noise as a matrix of values.

The pattern created below will use `ambient::noise_simple()` to create a particular 
type of noise.  This noise will be mapped to a colour gradient between two
user-specified colours.


Create the `ambient` pattern function
------------------------------------------------------------------------------

All array-based pattern creation functions must:

1. Have the exact function signature: `function(width, height, params, legend)`
2. Return a 3D RGBA numeric array with dimensions [height, width, 4]


Since `ambient` only creates a 2D matrix of noise, we will use a colour ramp to
promote each value in the matrix to create an RGBA value.


```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create an array of noise using the 'ambient' package
#'
#' @param width,height area dimensions
#' @param params aesthetic parameters passed from the geom e.g. 'pattern_fill', 
#'        'pattern_frequency' etc.
#' @param legend logical. If the request to create a pattern comes during 
#'        creation of the legend, then this is TRUE, otherwise FALSE
#'
#' @return an RGBA numeric array with dimensions [height, width, 4]
#'
#' @import ambient
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
create_pattern_ambient <- function(width, height, params, legend) {
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # The only 2 parameters needed are the 2 ends of the colour scale
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  colour1 <- as.character(params$pattern_fill )
  colour2 <- as.character(params$pattern_fill2)
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a ramp function from these 2 colours
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  ramp_func <- colorRamp(c(colour1, colour2), alpha = TRUE)
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a noise matrix of the requested dimenions using 'ambient'.
  # The contents are normalised to all be in the range [0,1]
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  noise_matrix  <- ambient::noise_simplex(dim = c(height, width))
  noise_matrix  <- ambient::normalise(noise_matrix)
  
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Use each value in the noise matrix to lookup a colour using the 
  # colour ramp function, then ensure the results are an RGBA array of the
  # correct dimensions.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  colour_matrix <- ramp_func(noise_matrix)/255
  noise_array   <- array(colour_matrix, dim = c(height, width, 4))
  
  
  noise_array
}
```


Let `{ggpattern}` know that there's an external pattern function it can use
------------------------------------------------------------------------------

A global option (`ggpattern_array_funcs`) is a named list which contains 
array creating functions to use outside of `ggpattern`.

The **name** used in this list corresponds to the `pattern` name used with 
the geom - in this case we will be using `pattern = 'ambient2'`.


```{r}
options(ggpattern_array_funcs = list(ambient2 = create_pattern_ambient))
```


Use this `ambient2` pattern
------------------------------------------------------------------------------

* ambient noise takes two colours which will be passed in via the parameters:
    * `pattern_fill` will be mapped to the `trt` variable
    * `pattern_fill2` will always be 'white'

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create some data to plot
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a ggplot using this pattern
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(pattern_fill = trt),
    pattern       = 'ambient', 
    pattern_fill2 = 'white',
    colour        = NA, 
    fill          = NA
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'ambient2'"
  ) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```



Mapping both colours simultaneously
------------------------------------------------------------------------------

```{r}
ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x = mpg, 
      pattern_fill  = as.factor(cyl), 
      pattern_fill2 = as.factor(cyl)
    ), 
    pattern = 'ambient2'
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(2, 'cm')) + 
  scale_pattern_fill_brewer (palette = 'Accent', direction =  1) + 
  scale_pattern_fill2_brewer(palette = 'Dark2' , direction =  1) + 
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'ambient2'"
  ) 
```



Future
------------------------------------------------------------------------------

* A more complete version of this pattern generator would need the following:
    * making all the different `ambient` noise types available (perhaps through `pattern_name`)
    * being able to configure the frequence, phase and other parameters for 
      each of the noise types.
* There is simple `plasma` pattern in `ggpattern` already, but the `ambient` 
  package has a lot more to offer.
* This might work better as a standalone package?  `ggpatternambient` ?














