---
title: "Experimental Pattern - Points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental Pattern - Points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6
)

suppressPackageStartupMessages({
  library(rlang)
  library(ggplot2)
  library(sf)
  library(grid)
  library(ggpattern)
})
```



Introduction to the geometry-based `hex` pattern
------------------------------------------------------------------------------


The pattern is an attempt to create a structured pattern of point elements using
only geometry elements.  The plotted points correspond to the plotting shapes used in 
`geom_point` and using `base` plotting.



Create the `points` pattern function
------------------------------------------------------------------------------

All geometry-based pattern creation functions must:

1. Have the exact function signature: `function(params, boundary_df, aspect_ratio, legend)`
2. Return a grid grob object



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create a pointsGrob object for a set of points
#'
#' Use 'sf' to help with the point in polygon intersections.
#'
#' \itemize{
#'   \item{make grid to cover entire space}
#'   \item{rotate points into position}
#'   \item{create expanded boundary by r}
#'   \item{create contracted boundary by r}
#'   \item{remove all points outside the expanded boundary}
#'   \item{remove all points within contracted boundary -> internal points}
#'   \item{any remaining points become part of the intersection grob}
#'   \item{total points = treeGrob( internal_internal, intersection_points)}
#' }
#'
#' @param boundary_df polygon_df data.frame
#' @param angle angle of orientation (degrees)
#' @param spacing spacing in grid 'npc' coordinates. Usually in range [0, 1]
#' @param density fill fraction. Number in range [0, 1]
#' @param xoffset,yoffset offset the pattern creation origin.
#' @param aspect_ratio aspect_ratio
#' @param params params from the geom
#'
#' @return A grid::pointsGrob
#'
#' @import grid
#' @import sf
#' @import grid
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
create_points_grob <- function(boundary_df, params, angle=0, spacing=0.1, density=0.3,
                                xoffset=0, yoffset=0,
                                aspect_ratio) {

  angle <- angle %% 90

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Calculate radius
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  r <- spacing * density / 2
  if (aspect_ratio > 1) {
    r <- r * aspect_ratio
  }

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Things get dicey at the boundaries, especially when there is very large
  # or small aspect ratio.   Include this fudge factor in buffering the
  # boundary to ensure that all partially ntersecting points are kept
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  fudge_factor <- aspect_ratio
  if (fudge_factor < 1) {
    fudge_factor <- 1/fudge_factor
  }


  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Generate a square grid of points
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  rff <- r * fudge_factor
  yff <- 1 / aspect_ratio * 1

  yff <- max(yff, 2)


  point_coords <- expand.grid(
    x = seq(-rff    , yff+rff, spacing),
    y = seq(-yff-rff, yff+rff, spacing)
  )

  if (nrow(point_coords) == 0) {
    return(grid::nullGrob())
  }

  point_coords   <- rotate_polygon_df(point_coords, angle, aspect_ratio)
  point_coords$y <- point_coords$y * aspect_ratio

  points_sf    <- sf::st_multipoint(as.matrix(point_coords))

  boundary_sf   <- convert_polygon_df_to_polygon_sf(boundary_df, buffer_dist =  0)
  expanded_sf   <- convert_polygon_df_to_polygon_sf(boundary_df, buffer_dist =  r * fudge_factor)
  contracted_sf <- convert_polygon_df_to_polygon_sf(boundary_df, buffer_dist = -r * fudge_factor)

  all_points_sf      <- sf::st_intersection(expanded_sf, points_sf)
  interior_points_sf <- sf::st_intersection(contracted_sf, all_points_sf)
  exterior_points_sf <- sf::st_difference(all_points_sf, contracted_sf)

  interior_points_mat <- as.matrix(interior_points_sf)
  exterior_points_mat <- as.matrix(exterior_points_sf)

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a grob for the internal points
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (is.null(interior_points_mat) || nrow(interior_points_mat) == 0) {
    interior_points_grob <- nullGrob()
  } else {
    interior_points_grob <- grid::pointsGrob(
      x    = interior_points_mat[,1],
      y    = interior_points_mat[,2],
      pch  = params$pattern_shape,
      size = unit(1, 'char'),
      gp   = gpar(
        fill = scales::alpha(params$pattern_fill  , params$pattern_alpha),
        col  = scales::alpha(params$pattern_colour, params$pattern_alpha),
        lwd  = params$pattern_size,
        lty  = params$pattern_linetype
      )
    )
  }

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create a grob for the intersecting points that don't lie
  # completely in the region.
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  if (is.null(exterior_points_mat) || nrow(exterior_points_mat) == 0) {
    exterior_points_grob <- nullGrob()
  } else {
    exterior_points_grob <- grid::pointsGrob(
      x    = exterior_points_mat[,1],
      y    = exterior_points_mat[,2],
      pch  = params$pattern_shape,
      size = unit(1, 'char'),
      gp   = gpar(
        fill = 'red', #scales::alpha(params$pattern_fill  , params$pattern_alpha),
        col  = 'green', #scales::alpha(params$pattern_colour, params$pattern_alpha),
        lwd  = params$pattern_size,
        lty  = params$pattern_linetype
      )
    )

  }


  grid::grobTree(
    interior_points_grob#,
    # exterior_points_grob
  )
}
```



```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
#' Create an array of noise using the 'ambient' package
#'
#' @param params aesthetic parameters passed from the geom e.g. 'pattern_fill', 
#'        'pattern_frequency' etc.
#' @param boundary_df is a data.frame of (x, y) coordinates of the boundary of
#"        the geom to be filled.
#' @param aspect_ratio this is the best guess of the current aspect ratio of the
#'        viewport into which the geometry is being drawn
#' @param legend logical. If the request to create a pattern comes during 
#'        creation of the legend, then this is TRUE, otherwise FALSE
#'
#' @return a grid grob object containing the pattern
#'
#' @import ambient
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
create_pattern_points <- function(params, boundary_df, aspect_ratio, legend = FALSE) {

  stopifnot(is_polygon_df(boundary_df))

  boundary_grob <- convert_polygon_df_to_polygon_grob(boundary_df)
  bbox          <- calculate_bbox_polygon_df(boundary_df)

  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  # Create an SF object with points covering the entire viewpoint
  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
  points_grob <- create_points_grob(
    boundary_df  = boundary_df,
    params       = params,
    angle        = params$pattern_angle,
    spacing      = params$pattern_spacing,
    density      = params$pattern_density,
    xoffset      = params$pattern_xoffset,
    yoffset      = params$pattern_yoffset,
    aspect_ratio = aspect_ratio
  )

  points_grob
}
```


Let `{ggpattern}` know that there's an external pattern function it can use
------------------------------------------------------------------------------

A global option (`ggpattern_geometry_funcs`) is a named list which contains 
geometry-based patten creating functions to use outside of `ggpattern`.

The **name** used in this list corresponds to the `pattern` name used with 
the geom - in this case we will be using `pattern = 'points'`.


```{r}
options(ggpattern_geometry_funcs = list(points = create_pattern_points))
```


Use this `points` pattern
------------------------------------------------------------------------------

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create some data to plot
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Create a ggplot using this pattern
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
p <- ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_shape  = trt, 
      pattern_angle  = trt, 
      pattern_colour = trt
    ), 
    pattern          = 'points',
    pattern_spacing  = 0.02,
    colour           = 'black', 
    pattern_density  = 0.15, 
    fill             = 'white',
    pattern_option_1 = 0.1
  ) +
  theme_bw() +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'points'"
  ) +
  theme(legend.position = 'none') +
  scale_pattern_angle_discrete(range = c(0, 30)) + 
  coord_fixed(ratio = 1/2)

p
```




Future
------------------------------------------------------------------------------

* Figure out some sizes of all the possible shapes so that the code can correctly 
  calculate the buffer around the boundary and ensure the shape doesn't overlap
* Manually create grid or `{sf}` polygon objects which recreate the shape - this 
  way we could use `gridGeometry::polyclipGrob` to clip the shapes at the boundary.
    * This looks like a lot of manual work.  Could perhaps raid the `{grid}` package
      and find the source for how the shapes are drawn and just clone that?
* All the drawn shapes are slightly different sizes, so it's difficult to write
  code to ensure that any drawn shape lies totally within the boundary.
* Legends are still funky
* Because points aren't polygons, `gridGeometry::polyclipGrob` will not clip
  a shape to a boundary.  so each shape is either drawn or not drawn, and there
  is currently no way to generate a partial shape that intersects the boundary.



