---
title: "Patterns 'magick' and 'polygon_tiling' - Parameters and Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Patterns 'magick' and 'polygon_tiling' - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(ggpattern)
})
```

## Introduction to the 'magick' and 'polygon_tiling' patterns

* The 'magick' pattern allows use of `imagemagick`'s in-built pixel-based patterns.
  It is an "array" pattern that depends on the suggested package `{magick}`.
* The 'polygon_tiling' pattern supports a large number of polygon tilings.
  It is a "geometry" pattern.

## `pattern = 'magick'` - filling a geom with pattern from `imagemagick`

`imagemagick` has the ability to create pixel-based patterns.  This pattern uses
those pixel patterns to fill the geom.

`pattern = 'magick'` options
------------------------------------------------------------------------------

| Aesthetic                  | Description                        | Default   |
|----------------------------|------------------------------------|-----------|
| `pattern_type`             | Image scaling type                 | 'fit'     |
| `pattern_fill`             | Colour                             | 'grey20'  |
| `pattern_scale`            | Extra scaling                      | 1         |
| `pattern_filter`           | Filter to use when scaling         | 'box    ' |
| `pattern_alpha`            | Alpha                              | NA        |
| `pattern_aspect_ratio`     | Override aspect ratio              | NA        |
| `pattern_key_scale_factor` | Additional scale factor for legend | 1         |


### `pattern = 'magick'` - `pattern_type` options

Patterns have been scaled (using `pattern_scale`) to make them more visible for this vignette.

Note that these are all pixel patterns and will always look blocky.

```{r fig.width=7, fig.height = 12, echo = TRUE}
if (require("magick")) {

df1 <- data.frame(
  x    = rep(1:6, 9),
  y    = rep(1:9, each=6),
  name = gridpattern::names_magick,
  stringsAsFactors = FALSE
)

ggplot(df1) + 
  geom_tile_pattern(
    aes(x, y, pattern_type = I(name)),
    pattern       = 'magick',
    pattern_scale = 1.5,
    pattern_fill  = 'white', 
    width         = 0.9, 
    height        = 0.9
  ) + 
  geom_label(aes(x+0.4, y+0.4, label = name), hjust = 1, vjust = 1) + 
  theme_void() + 
  labs(
    title = "All the possible magick pattern names"
  ) +
  coord_fixed(1)

}
```

### Example: `pattern = 'magick'` - Transparent patterns

```{r}
if (require("magick")) {

df1 <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))

ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_type = trt, 
      pattern_fill = trt
    ),
    pattern       = 'magick',
    pattern_key_scale_factor = 0.7,
    fill          = 'white',
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='magick'"
  ) +
  theme(legend.key.size = unit(2, 'cm')) +
  scale_pattern_type_discrete(choices = c('bricks', 'fishscales', 'right45')) +
  coord_fixed(ratio = 1/2)

}
```

### Example: `pattern = 'magick'` - Transparent patterns (2)

```{r}
if (require("magick")) {

ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x             = mpg,
      pattern_fill  = as.factor(cyl),
      pattern_type  = as.factor(cyl)
    ),
    pattern      = 'magick',
    pattern_scale = 2
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(2, 'cm')) +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern='magick'"
  ) +
  scale_pattern_type_discrete(choices = gridpattern::names_magick) +
  coord_fixed(ratio = 80)

}
```

### Example: `pattern = 'magick'` - Pie chart

```{r}
if (require("magick")) {

df2 <- data.frame(
  group = factor(c("Cool", "But", "Use", "Less"), levels = c("Cool", "But", "Use", "Less")),
  value = c(10, 20, 30, 40)
)

ggplot(df2, aes(x="", y = value, pattern_angle = group))+
  geom_bar_pattern(
    aes(pattern_type = group, pattern_fill = group),
    pattern                  = 'magick',
    pattern_scale            = 2,
    width                    = 1,
    stat                     = "identity",
    fill                     = 'white',
    colour                   = 'black',
    pattern_aspect_ratio     = 1,
    pattern_density          = 0.3
  ) +
  coord_polar("y", start=0) +
  theme_void(20) +
  theme(legend.key.size = unit(2, 'cm')) +
  scale_pattern_type_discrete(choices = gridpattern::names_magick_stripe) +
  labs(
    title    = "ggpattern::geom_bar_pattern()",
    subtitle = "pattern='magick'"
  ) 

}
```

## `pattern = 'polygon_tiling'` - filling a geom with a polygon tiling

### Pattern Parameters

| aesthetic                | description                                   | default   | possible values                   |
|--------------------------|-----------------------------------------------|-----------|-----------------------------------|
| pattern\_colour           | Stroke colour                                 | 'grey20'  | colour                            |
| pattern\_fill             | Fill colour                                   | 'grey80'  | colour                            |
| pattern\_angle            | Rotation angle                                | 30        | angle in degrees                  |
| pattern\_spacing          | Spacing between repetitions of pattern        | 0.05      | value in range [0, 1] (snpc units) |
| pattern\_xoffset          | Shift pattern along x axis                    | 0         | value in range [0, 1] (snpc units) |
| pattern\_yoffset          | Shift pattern along y axis                    | 0         | value in range [0, 1] (snpc units) |
| pattern\_alpha            | Alpha                                         | NA        | value in range [0, 1] or NA |
| pattern\_linetype         | Stroke linetype                               | 1         | linetype                          |
| pattern\_size             | Stroke linewidth                              | 1         | linewidth                         |
| pattern\_type             | Pattern type                                  | NA        | `gridpattern::names_polygon_tiling`.  See [polygon_tiling](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_polygon\_tiling.html) pattern documentation for more info. |

### `pattern = 'polygon_tiling'` - `pattern_type` options

```{r fig.width=7, fig.height = 12, echo = TRUE}

df1 <- data.frame(
  x    = rep(1:4, 9),
  y    = rep(1:9, each=4),
  name = gridpattern::names_polygon_tiling[1:36],
  stringsAsFactors = FALSE
)

ggplot(df1) + 
  geom_tile_pattern(
    aes(x, y, pattern_type = I(name)),
    pattern       = 'polygon_tiling',
    pattern_scale = 1.5,
    pattern_fill  = 'white', 
    width         = 0.9, 
    height        = 0.9
  ) + 
  geom_label(aes(x+0.4, y+0.4, label = name), hjust = 1, vjust = 1) + 
  theme_void() + 
  labs(
    title = "All the possible 'polygon_tiling' pattern names"
  ) 
```

### `pattern = 'polygon_tiling'` - Bar chart

```{r}
gg <- ggplot(mtcars) +
   geom_density_pattern(
     aes(
       x            = mpg,
       pattern_fill = as.factor(cyl),
       pattern_type = as.factor(cyl)
     ),
     pattern = 'polygon_tiling',
     pattern_key_scale_factor = 1.2
   ) +
   scale_pattern_type_manual(values = c("hexagonal", "rhombille",
                              "pythagorean")) +
   theme_bw(18) +
   theme(legend.key.size = unit(2, 'cm')) +
   labs(
     title    = "ggpattern::geom_density_pattern()",
     subtitle = "pattern = 'polygon_tiling'"
   )
plot(gg)
```


