---
title: "Patterns 'gradient', 'ambient', 'plasma' - Parameters and Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Patterns 'gradient', 'ambient', 'plasma' - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(ggpattern)
})
```

## Introduction to the 'gradient', 'ambient', and 'plasma' patterns

* The 'gradient' pattern fills the geom with color gradients.
  This pattern either depends on the R package `{magick}`
  or a graphics device and version of R that supports the
  new R 4.1 gradient feature.
* The 'plasma' pattern uses Imagemagick (`magick`) to generate some 
  fractal noise arrays which are coloured and used to fill the geom.
  This pattern depends on the R package `{magick}`.
* The `{ambient}` package offers a way of creating multiple different types of 
  noise as a matrix of values.  The 'ambient' pattern maps this noise 
  to a colour gradient between two user-specified colours.

## Pattern Parameters

| aesthetic                  | description                           | default   |
|----------------------------|---------------------------------------|-----------|
| `pattern_fill`             | Noise colour                          | 'grey80'  |
| `pattern_fill2`            | Second colour ('ambient', 'gradient') | '#4169E1' |
| `pattern_scale`            | Extra scaling                         | 1         |
| `pattern_alpha`            | Alpha                                 | NA        |
| `pattern_type`             | Type of noise ('ambient')             | 'simplex' |
| `pattern_frequency`        | Granularity of features in noise      | 0.1       |
| `pattern_aspect_ratio`     | Override aspect ratio                 | NA        |
| `pattern_key_scale_factor` | Additional scale factor for legend    | 1         |

## Example Data

```{r}
df <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2)
)
```

## Example: `pattern = 'plasma'` - With `pattern_alpha`

```{r}
if (require("magick")) {

ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill         = trt,
      pattern_fill = trt
    ),
    fill          = NA,
    pattern       = 'plasma',
    pattern_alpha = 1,
    pattern_scale = 2,
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='plasma'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm'))

}
```

```{r}
if (require("magick")) {

ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x             = mpg,
      pattern_fill  = as.factor(cyl)
    ),
    pattern      = 'plasma',
    pattern_alpha = 0.7
  ) +
  theme_bw(15) +
  theme(legend.position = 'none') +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern='plasma'"
  )

}
```

Example: `pattern_type = 'gradient'` - Fade to white
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_fill        = trt,
      pattern_orientation = trt
    ), 
    pattern       = 'gradient', 
    pattern_fill2 = 'white', 
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'gradient'"
  ) +
  theme(
    legend.key.size = unit(1.5, 'cm')
  )

}
```


Example: `pattern_type = 'gradient'` - Fade to dark blue
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_fill        = trt,
      pattern_orientation = trt
    ), 
    pattern       = 'gradient', 
    pattern_fill2 = '#445566', 
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'gradient'"
  ) +
  theme(
    legend.key.size = unit(1.5, 'cm')
  )

}
```


Example: `pattern_type = 'gradient'` - Fade to transparent
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_fill        = trt,
      pattern_orientation = trt
    ), 
    pattern       = 'gradient', 
    pattern_fill2 = NA, 
    fill          = NA,
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'gradient'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm'))

}
```

## Example: `pattern_type = 'gradient'` - Non-rectangular geom

```{r}
if (require("magick")) {

ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x = mpg, 
      pattern_fill = as.factor(cyl),
      pattern_orientation = as.factor(cyl)
    ), 
    pattern       = 'gradient',
    pattern_fill2 = NA, 
    fill          = NA
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'gradient'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm'))

}
```

## Example: `pattern_type = 'ambient'

```{r}
if (require("ambient")) {
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(pattern_fill = trt),
    pattern       = 'ambient', 
    pattern_fill2 = 'white',
    colour        = NA, 
    fill          = NA
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'ambient'"
  ) +
  theme(legend.position = 'none')
}
```

```{r}
if (require("ambient")) {
ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x = mpg, 
      pattern_fill  = as.factor(cyl), 
      pattern_fill2 = as.factor(cyl)
    ), 
    pattern = 'ambient'
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(2, 'cm')) + 
  scale_pattern_fill_brewer (palette = 'Accent', direction =  1) + 
  scale_pattern_fill2_brewer(palette = 'Dark2' , direction =  1) + 
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'ambient'"
  ) 
}
```
