---
title: "Pattern Image - Parameters and Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pattern Image - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width  = 7,
  fig.height = 5
)
```

```{r setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(ggpattern)
})
```


The `image` pattern - filling a geom with an image from file or URL
------------------------------------------------------------------------------

This pattern will load an image from a file or URL  (`pattern_filename`)
and display it within the geom.

Because the image will most likely not exactly cover the area of the geom, there
are options to specify how the image is used to fill 
the region.

After expanding to cover the area, the image is then clipped to the boundary of the geom.

The files/URLs are loaded with `magick::image_read` which means that transparent
images and SVGs are supported.

`image` options
------------------------------------------------------------------------------

| Aesthetic                  | Description                        | Default   |
|----------------------------|------------------------------------|-----------|
| `pattern_filename`         | Image filename or URL              | ''        |
| `pattern_type`             | Image scaling type                 | 'fit'     |
| `pattern_scale`            | Extra scaling                      | 1         |
| `pattern_gravity`          | Position of image within area      | 'center'  |
| `pattern_filter`           | Filter to use when scaling         | 'lanczos' |
| `pattern_alpha`            | Alpha                              | NA        |
| `pattern_aspect_ratio`     | Override aspect ratio              | NA        |
| `pattern_key_scale_factor` | Additional scale factor for legend | 1         |

| `pattern_type` | Description                                                                                                                     | `pattern_scale` | `pattern_gravity`           |
|----------------|---------------------------------------------------------------------------------------------------------------------------------|-----------------|-----------------------------|
| squish         | distort the image to cover the bounding box of the region                                                                       |                 |                             |
| fit            | scale the image such that either the width or the height of the image    fits in the bounding box.                              |                 | Yes                         |
| expand         | scale the image beyond the bounding box and crop it such that the     image fully covers the width and the height of the region |                 |                             |
| none           | position a single image in the region without attempting to scale to the bounding box size                                      | Yes             | Yes                         |
| tile           | repeat the image to cover the bounding box.                                                                                     | Yes             |                             |


Note: not all combinations of aesthetics are useful e.g. if `pattern_type = 'fit'` is specified, then
`pattern_scale` has no effect.


Example Data
------------------------------------------------------------------------------

These examples use some of the built-in images in the `ggpattern` package.

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# filenames of images
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
logo_filename   <- system.file("img", "Rlogo.png" , package="png")
magpie_filename <- system.file("img", "magpie.jpg", package="ggpattern")
bug_filename    <- system.file("img", "bug.jpg"   , package="ggpattern")

seamless1 <- system.file("img", "seamless1.jpg"   , package="ggpattern")
seamless2 <- system.file("img", "seamless2.jpg"   , package="ggpattern")
seamless3 <- system.file("img", "seamless3.jpg"   , package="ggpattern")

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Plotting data
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df1 <- tibble(
  trt      = c("a", "b", "c"), 
  outcome  = c(2.3, 1.9, 3.2),
  gravity  = c('South', 'North', 'West'),
  filltype = c('squish', 'fit' , 'expand'),
  scale    = c(1, 2, 0.5),
  filename = c(logo_filename, magpie_filename, bug_filename)
)
```

```{r echo = FALSE}
knitr::kable(df1, caption = "Example Data")
```




Example: `pattern = 'image'` - No Attempt at filling the area (`pattern_type = 'none'`)
------------------------------------------------------------------------------

If `pattern_type = 'none'` then no attempt is made at snugly filling the area.
Instead a single copy of the image is scaled (`pattern_scale`) and 
placed (`pattern_gravity`) in the geom area.


```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill            = trt,
      pattern_gravity = I(gravity),
      pattern_scale   = I(scale)
    ), 
    pattern          = 'image', 
    pattern_filename = logo_filename,
    pattern_type     = 'none',
    colour           = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'image', pattern_type = 'none'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm')) +
  coord_fixed(ratio = 1/2)
```


Example: `pattern = 'image'` - Covering the area (`pattern_type = 'fit', 'squish', 'expand'`)
------------------------------------------------------------------------------

In the plot below, 3 different methods are used to cover the area of the geom. From 
left-to-right:

* `pattern_type = 'squish'` to force the image to cover the bounding box of the geom.
* `pattern_type = 'fit'` to force the image to cover the width (or height) yet still
  remain fully visible.  `pattern_gravity = 'North'` floats the image to the top.
* `pattern_type = 'expand`` scales the image (while preserving aspect ratio) such 
  that the entire geom is covered.  `pattern_gravity = 'West'` anchors the image
  to the left side of the geom.

```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill            = trt,
      pattern_gravity = I(gravity),
      pattern_type    = I(filltype)
    ), 
    pattern          = 'image', 
    colour           = 'black',
    pattern_filename = logo_filename,
    
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'image'"
  ) +
  # theme(legend.key.size = unit(1.5, 'cm')) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```



Example: `pattern = 'image'` - Tiling (`pattern_type = 'tile'`)
------------------------------------------------------------------------------

When tiled, the image is replicated in order to cover the entire area of the geom.
In this example, `pattern_scale` is also applied to the image so that it appears at
different sizes.


```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill            = trt,
      pattern_gravity = I(gravity),
      pattern_scale   = I(scale)
    ), 
    pattern          = 'image', 
    pattern_type     = 'tile',
    pattern_filename = logo_filename,
    colour           = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'image', pattern_type = 'tile'"
  ) +
  # theme(legend.key.size = unit(1.5, 'cm')) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```




Example: `pattern = 'image'` - Tiling with seamless patterns
------------------------------------------------------------------------------

By choosing a seamlessly tiling image, then a tiled fill will not have visible discontinuities.

```{r}
ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x = mpg, 
      pattern_filename = as.factor(cyl)
    ), 
    pattern      = 'image',
    pattern_type = 'tile'
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(1.5, 'cm')) + 
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'image', pattern_type = 'tile'"
  ) +
  scale_pattern_filename_manual(values = c(`4` = seamless1, `6` = seamless2, `8` = seamless3)) +
  coord_fixed(ratio = 80)
```



Example: `pattern = 'image'` - Tiling with seamless patterns (with scaling)
------------------------------------------------------------------------------

```{r}
ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x = mpg, 
      pattern_filename = as.factor(cyl)
    ), 
    pattern       = 'image',
    pattern_type  = 'tile',
    pattern_scale = 0.5
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(2, 'cm')) + 
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern = 'image', pattern_type = 'tile'"
  ) +
  scale_pattern_filename_manual(values = c(`4` = seamless1, `6` = seamless2, `8` = seamless3)) +
  coord_fixed(ratio = 80)
```



Example: `pattern = 'image'` - Tiling with fit to width (of element bounding box)
------------------------------------------------------------------------------

In a common case, the image to be tiled should be first scaled to
the width of the bar to be tiled and then tiling appears like a vertical stacking.
This is useful for bar charts.

Specify `pattern_scale = -1` to fit the width of the geom, and `pattern_scale = -2` to 
fit the height of the geom.

```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      fill             = trt,
      pattern_filename = I(filename)
    ), 
    pattern          = 'image', 
    pattern_type     = 'tile',
    pattern_scale    = -1,
    colour           = 'black'
    
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern = 'image', pattern_type = 'tile',\npattern_scale = -1"
  ) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1)
```

