---
title: "Patterns 'circle', 'pch', 'regular_polygon', 'rose' - Parameters and Examples"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Patterns 'circle', 'pch', 'regular_polygon', 'rose' - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, 
  fig.height = 4
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(ggpattern)
})
```

## Introduction to the 'circle', 'pch', 'regular_polygon', and 'rose' patterns

* The 'circle' pattern draws circles
* The 'pch' pattern draws "plotting character" symbols as in `graphics::points()`
* The 'regular\_polygon' pattern draws convex/star regular polygons
* The 'rose' pattern draws [rose curves](https://en.wikipedia.org/wiki/Rose\_(mathematics))

## Pattern Parameters

| aesthetic                 | description                                   | default   | possible values                   |
|---------------------------|-----------------------------------------------|-----------|-----------------------------------|
| pattern\_colour           | Stroke colour                                 | 'grey20'  | colour                            |
| pattern\_fill             | Fill colour                                   | 'grey80'  | colour                            |
| pattern\_angle            | Rotation angle                                | 30        | angle in degrees                  |
| pattern\_density          | Approx. fraction of area the pattern fills    | 0.2       | value in range [0, 1] (fraction)  |
| pattern\_spacing          | Spacing between repetitions of pattern        | 0.05      | value in range [0, 1] (snpc units) |
| pattern\_xoffset          | Shift pattern along x axis                    | 0         | value in range [0, 1] (snpc units) |
| pattern\_yoffset          | Shift pattern along y axis                    | 0         | value in range [0, 1] (snpc units) |
| pattern\_alpha            | Alpha                                         | NA        | value in range [0, 1] or NA |
| pattern\_linetype         | Stroke linetype                               | 1         | linetype                          |
| pattern\_size             | Stroke linewidth                              | 1         | linewidth                         |
| pattern\_type             | Pattern type                                  | NA        | See [circle](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_circle.html), [pch](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_pch.html), [regular\_polygon](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_regular\_polygon.html), and [rose](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_rose.html) pattern documentation |
| pattern\_subtype          | Pattern subtype                               | NA        | See [circle](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_circle.html), [pch](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_pch.html), [regular\_polygon](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_regular\_polygon.html), and [rose](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_rose.html) pattern documentation |
| pattern\_grid             | Grid type                                     | 'square'  | 'square', 'hex', or 'hex\_circle' |
| pattern\_rot              | Pattern symbol rotation angle                 | 0         | angle in degrees                  |
| pattern\_shape            | Pattern shape                                 | NA        | See [pch](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_pch.html) and [regular\_polygon](https://trevorldavis.com/R/gridpattern/dev/reference/grid.pattern\_regular\_polygon.html) |
| pattern\_frequency        | Frequency                                     | 0.1       | Angular frequency in 'rose' pattern |

There are also a number of parameters for extra control of legend sizing and
aspect ratio adjustments. See the 'Pattern Parameters - Common' for more 
information.

## Data

Standard data for all the example plots

```{r}
df <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))
df
```

## Vanilla ggplot2

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col(aes(fill=trt),colour='black') +
  theme_bw() +
  labs(title = "Plain ggplot2")
```

## Use the `{ggpattern}` geom

* Use `ggpattern::geom_col_pattern()` instead of `ggplot2::geom_col()`.
* Set `pattern = 'circle'`
* Default legends in `ggplot2` are usually too small to show off an example of
  the pattern, so it will usually be necessary to increase the key size.

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(aes(fill=trt),colour='black', pattern = 'circle') +
  theme_bw() +
  labs(title = "ggpattern") +
  theme(legend.key.size = unit(1.5, 'cm'))
```

## The Density Aesthetic ('pch' example)

The aesthetic `pattern_density` roughly corresponds to the fraction of the filled
area which should be covered by the pattern.

In the following plot the density of striping is increased to 50% of the fill area.

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill=trt, pattern_shape=trt),
    colour          = 'black', 
    pattern         = 'pch',
    pattern_density = 0.5
  ) +
  theme_bw() +
  labs(title = "Fixed density of 0.5 (50% of the fill area)") + 
  theme(legend.key.size = unit(1.5, 'cm'))
```

## The Density Aesthetic as a Mapped Aesthetic

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_density = trt),
    colour          = 'black', 
    pattern         = 'circle'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Density") + 
  theme(legend.key.size = unit(1.5, 'cm'))
```

## The Density Aesthetic as a Mapped Aesthetic with Manual Scale

`scale_pattern_density_manual()` can be used to manually control how the variable 
is mapped to the density.

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_density = trt),
    colour          = 'black', 
    pattern         = 'regular_polygon',
    pattern_shape   = 'convex6',
    pattern_fill    = 'white',
    pattern_colour  = NA,
    pattern_grid    = 'hex'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Density") + 
  theme(legend.key.size = unit(1.5, 'cm')) + 
  scale_pattern_density_manual(values = c(a = 0.7, b=0.8, c=0.9))
```

## The Spacing Aesthetic as a Mapped Aesthetic

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_spacing = trt, pattern_shape = trt),
    pattern_density = 0.7,
    colour          = 'black', 
    pattern         = 'regular_polygon',
    pattern_scale   = 0.2
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Spacing") + 
  scale_pattern_shape_manual(values = c(a = "circle", b = "convex3", c = "star5")) +
  theme(legend.key.size = unit(1.5, 'cm'))
```

## The Fill Aesthetic as a Mapped Aesthetic ('rose' example)

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_fill = trt),
    colour          = 'black', 
    pattern         = 'rose',
    pattern_frequency = 5/2,
    pattern_density = 0.7,
    pattern_spacing = 0.15
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Pattern Fill") + 
  scale_pattern_fill_viridis_d() + 
  theme(legend.key.size = unit(1.5, 'cm'))
```
