---
title: "Pattern Placeholder - Parameters and Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pattern Placeholder - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.width = 6,
fig.height = 5
)
```

```{r setup}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(ggpattern)
})
```


`placeholder` - Filling with an image placeholder
------------------------------------------------------------------------------

Array-based patterns allow the user to specify an RGBA that should be displayed 
in the geom.

Getting the correct sized image such that it takes up the full space without 
being distorted can be a time-consuming task.

The **placeholder** pattern takes out this drudgery by fetching exactly the correct
sized image to fit the space.  These images come from image placeholder sites
which are most often used by web-developers as a stand-in for a final image 
while they are developing a new site.


`placeholder` options
------------------------------------------------------------------------------

| Aesthetic                  | Description                        | Default   |
|----------------------------|------------------------------------|-----------|
| `pattern_type`             | Image source                       | 'kitten'  |
| `pattern_alpha`            | Alpha                              | NA        |
| `pattern_aspect_ratio`     | Override aspect ratio              | NA        |
| `pattern_key_scale_factor` | Additional scale factor for legend | 1         |


`placeholder` option `pattern_type`
------------------------------------------------------------------------------

The following is a list of all the `pattern_type` values which are valid. If
unspecified or unknown, then `ggpattern` will use `pattern_type = 'kitten'`.

If you would like only greyscale images, append `bw` to the name, e.g. to 
display only black and white kittens use `pattern_type = 'kittenbw'`.

Click the link to see an example of each placeholder generator

* 'kitten'         - [kitten](https://placekitten.com/400/300)
* 'picsum'         - [random picture from picsum](https://picsum.photos/400/300)
* 'murray'         - [bill murray](http://www.fillmurray.com/400/300)
* 'cage'           - [nicholas cage](https://www.placecage.com/400/300)
* 'bear'           - [bears](https://placebear.com/400/300)
* 'seagal'         - [steven seagal](https://www.stevensegallery.com/400/300)
* 'placeholder'    - [placeholder.com](https://via.placeholder.com/400x300.png)
* 'dummy'          - [image showing dimensions](https://dummyimage.com/400x300)
* 'flickr'         - [random picture from flickr](https://loremflickr.com/400/300)
* 'beard'          - [beards](http://placebeard.it/400/300)
* 'lorempixel'     - [lorempixel](http://lorempixel.com/400/300/)
* 'placeimg'       - [random pic from placeimg.com](http://placeimg.com/400/300/any)
* 'keanu'          - [Keanu](https://placekeanu.com/400/300)

All placeholder names are available in `ggpattern::placeholder_names`.


Example Data
------------------------------------------------------------------------------

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Simple testing data
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df1 <- data.frame(
  trt     = c("a", "b", "c"), 
  outcome = c(2.3, 1.9, 3.2),
  stringsAsFactors = FALSE
)
```

Example: `pattern = 'placeholder'` - `pattern_type = 'bear'` 
------------------------------------------------------------------------------

```{r}

ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt), 
    pattern      = 'placeholder', 
    pattern_type = 'bear',
    colour       = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='placeholder', pattern_type='bear'"
  ) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```



Example: `pattern = 'placeholder'` - `pattern_type = 'bear'` with `pattern_alpha`
------------------------------------------------------------------------------

```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt), 
    pattern       = 'placeholder', 
    pattern_type  = 'bear',
    pattern_alpha = 0.5,
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='placeholder', pattern_type='bear'"
  ) +
  theme(legend.key.size = unit(1.5, 'cm')) +
  coord_fixed(ratio = 1/2)
```



Example: `pattern = 'placeholder'` - `pattern_type = 'picsum'`  
------------------------------------------------------------------------------

```{r}
ggplot(mtcars) +
  geom_density_pattern(
    aes(x = mpg, group = as.factor(cyl)),
    pattern      = 'placeholder',
    pattern_type = 'picsum'
  ) +
  theme_bw(15) +
  theme(legend.position = 'none') +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='placeholder', pattern_type='picsum'"
  ) +
  coord_fixed(ratio = 80)
```




Example: `pattern = 'placeholder'` - `pattern_type = 'dummy'` 
------------------------------------------------------------------------------



```{r}
ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt), 
    pattern       = 'placeholder', 
    pattern_type  = 'dummy',
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='placeholder', pattern_type='dummy'"
  ) +
  theme(legend.position = 'none') +
  coord_fixed(ratio = 1/2)
```



Example: `pattern = 'placeholder'` - `pattern_type = 'kitten'` 
------------------------------------------------------------------------------

```{r}
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Example data for pie chart
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
df2 <- data.frame(
  group = factor(c("Cool", "But", "Use", "Less"), levels = c("Cool", "But", "Use", "Less")),
  value = c(10, 20, 30, 40)
)

ggplot(df2, aes(x="", y = value, pattern_angle = group))+
  geom_bar_pattern(
    pattern      = 'placeholder',
    pattern_type = 'kitten',
    pattern_aspect_ratio = 1,
    width        = 1,
    stat         = "identity",
    colour       = 'white',
    size         = 2
  ) +
  coord_polar("y", start=0) +
  theme_void(15) +
  theme(legend.position = 'none') +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='placeholder', pattern_type='kitten'"
  ) 
```
