---
title: "Pattern Magick - Parameters and Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pattern Magick - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  fig.width = 8,
  fig.height = 6
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(ggpattern)
})
```


`pattern = 'magick'` - filling a geom with pattern from `imagemagick`
------------------------------------------------------------------------------

`imagemagick` has the ability to create pixel-based patterns.  This pattern uses
those pixel patterns to fill the geom.



`pattern = 'magick'` options
------------------------------------------------------------------------------

| Aesthetic                  | Description                        | Default   |
|----------------------------|------------------------------------|-----------|
| `pattern_type`             | Image scaling type                 | 'fit'     |
| `pattern_fill`             | Colour                             | 'grey20'  |
| `pattern_scale`            | Extra scaling                      | 1         |
| `pattern_filter`           | Filter to use when scaling         | 'box    ' |
| `pattern_alpha`            | Alpha                              | NA        |
| `pattern_aspect_ratio`     | Override aspect ratio              | NA        |
| `pattern_key_scale_factor` | Additional scale factor for legend | 1         |


`pattern = 'magick'` - `pattern_type` options
------------------------------------------------------------------------------

Patterns have been scaled (using `pattern_scale`) to make them more visible for this vignette.

Note that these are all pixel patterns and will always look blocky.

```{r fig.width=7, fig.height = 12, echo = TRUE}
if (require("magick")) {

df1 <- data.frame(
  x    = rep(1:6, 9),
  y    = rep(1:9, each=6),
  name = gridpattern::names_magick,
  stringsAsFactors = FALSE
)

ggplot(df1) + 
  geom_tile_pattern(
    aes(x, y, pattern_type = I(name)),
    pattern       = 'magick',
    pattern_scale = 1.5,
    pattern_fill  = 'white', 
    width         = 0.9, 
    height        = 0.9
  ) + 
  geom_label(aes(x+0.4, y+0.4, label = name), hjust = 1, vjust = 1) + 
  theme_void() + 
  labs(
    title = "All the possible magick pattern names"
  ) +
  coord_fixed(1)

}
```

Example: `pattern = 'magick'` - Transparent patterns
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

df1 <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))

ggplot(df1, aes(trt, outcome)) +
  geom_col_pattern(
    aes(
      pattern_type = trt, 
      pattern_fill = trt
    ),
    pattern       = 'magick',
    pattern_key_scale_factor = 0.7,
    fill          = 'white',
    colour        = 'black'
  ) +
  theme_bw(15) +
  labs(
    title    = "ggpattern::geom_col_pattern()",
    subtitle = "pattern='magick'"
  ) +
  theme(legend.key.size = unit(2, 'cm')) +
  scale_pattern_type_discrete(choices = c('bricks', 'fishscales', 'right45')) +
  coord_fixed(ratio = 1/2)

}
```


Example: `pattern = 'magick'` - Transparent patterns (2)
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

ggplot(mtcars) +
  geom_density_pattern(
    aes(
      x             = mpg,
      pattern_fill  = as.factor(cyl),
      pattern_type  = as.factor(cyl)
    ),
    pattern      = 'magick',
    pattern_scale = 2
  ) +
  theme_bw(15) +
  theme(legend.key.size = unit(2, 'cm')) +
  labs(
    title    = "ggpattern::geom_density_pattern()",
    subtitle = "pattern='magick'"
  ) +
  scale_pattern_type_discrete(choices = gridpattern::names_magick) +
  coord_fixed(ratio = 80)

}
```

Example: `pattern = 'magick'` - Pie chart
------------------------------------------------------------------------------

```{r}
if (require("magick")) {

df2 <- data.frame(
  group = factor(c("Cool", "But", "Use", "Less"), levels = c("Cool", "But", "Use", "Less")),
  value = c(10, 20, 30, 40)
)

ggplot(df2, aes(x="", y = value, pattern_angle = group))+
  geom_bar_pattern(
    aes(pattern_type = group, pattern_fill = group),
    pattern                  = 'magick',
    pattern_scale            = 2,
    width                    = 1,
    stat                     = "identity",
    fill                     = 'white',
    colour                   = 'black',
    pattern_aspect_ratio     = 1,
    pattern_density          = 0.3
  ) +
  coord_polar("y", start=0) +
  theme_void(20) +
  theme(legend.key.size = unit(2, 'cm')) +
  scale_pattern_type_discrete(choices = gridpattern::names_magick_stripe) +
  labs(
    title    = "ggpattern::geom_bar_pattern()",
    subtitle = "pattern='magick'"
  ) 

}
```
