---
title: "Pattern Stripe - Parameters and Examples"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Pattern Stripe - Parameters and Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7, 
  fig.height = 4
)
```

```{r setup}
suppressPackageStartupMessages({
  library(ggplot2)
  library(ggpattern)
})
```

## Introduction `pattern = 'stripe'`

The stripe pattern

## Pattern Parameters

| aesthetic                | description                                   | default   | possible values                   |
|--------------------------|-----------------------------------------------|-----------|-----------------------------------|
| pattern_colour           | Stroke colour                                 | 'grey20'  | colour                            |
| pattern_fill             | Fill colour                                   | 'grey80'  | colour                            |
| pattern_angle            | Rotation angle                                | 30        | angle in degrees                  |
| pattern_density          | Approx. fraction of area the pattern fills    | 0.2       | value in range [0, 1] (fraction)  |
| pattern_spacing          | Spacing between repetitions of pattern        | 0.05      | value in range [0, 1] (npc units) |
| pattern_xoffset          | Shift pattern along x axis                    | 0         | value in range [0, 1] (npc units) |
| pattern_yoffset          | Shift pattern along y axis                    | 0         | value in range [0, 1] (npc units) |
| pattern_alpha            | Alpha                                         | 1         | value in range [0, 1] (npc units) |
| pattern_linetype         | Stroke linetype                               | 1         | linetype                          |
| pattern_size             | Stroke linewidth                              | 1         | linewidth                         |


There are also a number of parameters for extra control of legend sizing and
aspect ratio adjustments. See the 'Pattern Parameters - Common' for more 
information.


## Data

Standard data for all the example plots

```{r}
df <- data.frame(trt = c("a", "b", "c"), outcome = c(2.3, 1.9, 3.2))
df
```


## Vanilla ggplot2

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col(aes(fill=trt),colour='black') +
  theme_bw() +
  labs(title = "Plain ggplot2")
```



## Use the `{ggpattern}` geom

* Use `ggpattern::geom_col_pattern()` instead of `ggplot2::geom_col()`.
* Set `pattern = 'stripe'`


```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(aes(fill=trt),colour='black', pattern = 'stripe') +
  theme_bw() +
  labs(title = "ggpattern")
```


## Use `coord_fixed()` 

Due to the underlying nature of `grid` graphics and how `ggplot2` sets up a plot, 
it is essential to use `coord_fixed()` in order to achieve consistent looking 
patterns.   Without a fixed coordinate system, then `ggpattern` has to make 
guesses on some sizes, and patterns may look askew e.g. 90 degree angles in patterns
may no longer appear properly!


```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(aes(fill=trt),colour='black', pattern = 'stripe') +
  theme_bw() +
  labs(title = "ggpattern + coord_fixed()") + 
  coord_fixed(ratio = 1/2)
```


## Bigger Legend Keys 

Default legends in `ggplot2` are usually too small to show off an example of
the pattern, so it will usually be necessary to increase the key size.


```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(aes(fill=trt),colour='black', pattern = 'stripe') +
  theme_bw() +
  labs(title = "ggpattern + coord_fixed()") + 
  coord_fixed(ratio = 1/2) + 
  theme(legend.key.size = unit(1.5, 'cm'))
```


## The Density Aesthetic

The aesthetic `pattern_density` roughly corresponds to the fraction of the filled
area which should be covered by the pattern.

In the following plot the density of striping is increased to 50% of the fill area.

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill=trt),
    colour          = 'black', 
    pattern         = 'stripe',
    pattern_density = 0.5
  ) +
  theme_bw() +
  labs(title = "Fixed density of 0.5 (50% of the fill area)") + 
  coord_fixed(ratio = 1/2) + 
  theme(legend.key.size = unit(1.5, 'cm'))
```


## The Density Aesthetic as a Mapped Aesthetic

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_density = trt),
    colour          = 'black', 
    pattern         = 'stripe'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Density") + 
  coord_fixed(ratio = 1/2) + 
  theme(legend.key.size = unit(1.5, 'cm'))
```


## The Density Aesthetic as a Mapped Aesthetic with Manual Scale


`scale_pattern_density_manual()` can be used to manually control how the variable 
is mapped to the density.


```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_density = trt),
    colour          = 'black', 
    pattern         = 'stripe'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Density") + 
  coord_fixed(ratio = 1/2) + 
  theme(legend.key.size = unit(1.5, 'cm')) + 
  scale_pattern_density_manual(values = c(a = 0.1, b=0.3, c=0.5))
```


## The Spacing Aesthetic as a Mapped Aesthetic

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_spacing = trt),
    colour          = 'black', 
    pattern         = 'stripe'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Spacing") + 
  coord_fixed(ratio = 1/2) + 
  theme(legend.key.size = unit(1.5, 'cm'))
```


## The Fill Aesthetic as a Mapped Aesthetic

```{r}
ggplot(df, aes(trt, outcome)) +
  geom_col_pattern(
    aes(fill = trt, pattern_fill = trt),
    colour          = 'black', 
    pattern         = 'stripe'
  ) +
  theme_bw() +
  labs(title = "Aesthetic Mapping of 'trt' to Pattern Fill") + 
  coord_fixed(ratio = 1/2) + 
  scale_pattern_fill_viridis_d() + 
  theme(legend.key.size = unit(1.5, 'cm'))
```




## Technical Bits

* Stripes are implemented as long thin polygons which are clipped to the boundaries of 
  the elements of the geom.
* The clipping of the stripe polygons is done by creating the stripes as `{sf}` objects, 
  performing an `sf::st_intersection()` with the element boundary, and then converting 
  the clipped stripes back into a `grid::polygonGrob()`.
* If lines had been used for the stripes, then there would be visual artefacts 
  where the stripe intersects the boundary because the `lineend` graphic parameter
  can only be `square`, `round` etc, and cannot conform to the actual shape of the 
  element boundary.
